---
title: "PA1_template"
author: "Ira Gu"
date: "April 6, 2018"
output: html_document
---

#Loading and preprocessing the data

```{r}
dat <- read.csv('./data/activity.csv', sep  = "," )
summary(dat)
```

#What is mean total number of steps taken per day?

Total step for every day

```{r}
dat1 <- aggregate(dat$steps, by = list(dat$date), FUN = sum, na.rm = TRUE)
names(dat1) <- c('date', 'totalsteps')
head(dat1, 3)
```

Histogram of total number of steps taken per day

```{r}
hist(dat1$totalsteps,
     main = "Steps Per Day",
     xlab = "Steps",
     ylab = "Days",
     col = "red")
```

Mean and Median steps per day

```{r}
median(dat1$totalsteps)
mean(dat1$totalsteps)
```

#What is the average daily activity pattern?

Average step for every 5 minute interval

```{r}
dat2 <- aggregate(dat$steps, by = list(dat$interval), FUN = mean, na.rm = TRUE)
names(dat2) <- c('interval', 'avg_steps')
head(dat2,3)
```

Plot of average daily activity pattern

```{r}
plot(dat2$interval, dat2$avg_steps, type = "l", xlab = 'interval', ylab = "avg_Steps")
```

5 Min interval with maximum steps

```{r}
dat2[which.max(dat2$avg_steps),]
```


#Imputing missing values

Total # of NA in each column

```{r}
sapply(dat, function(y) sum(is.na(y)))
```

New dataset where steps with value NA is replaced by the 5 min mean across all observations

```{r}
datf <- dat
datf$steps[is.na(datf$steps)] <- mean(dat$steps,na.rm= TRUE)
sapply(datf, function(y) sum(is.na(y)))
head(datf,3)
```

Create 2 datasets, 1 aggregate by mean and 1 aggregate by median 

```{r}
datf1 <- aggregate(datf$steps, by = list(datf$date), FUN = mean, na.rm = TRUE)
names(datf1) <- c('day', 'mean_steps')
head(datf1)
datf2 <- aggregate(datf$steps, by = list(datf$date), FUN = median, na.rm = TRUE)
names(datf2) <- c('day', 'median_Steps')
head(datf2)
```

Histogram for mean total steps per day

```{r}
hist(datf1$mean_steps,
         main = "Mean Steps Per Day",
         xlab = "Steps",
         ylab = "Days",
         col = "red")
```

Histogram for median total steps per day

```{r}
hist(datf2$median_Steps,
         main = "Median Steps Per Day",
         xlab = "Steps",
         ylab = "Days",
         col = "red")
```

Mean and median are both increased with additional inputs

```{r}
datf <- aggregate(datf$steps, by = list(datf$date), FUN = sum, na.rm = TRUE)
names(datf) <- c('date', 'totalsteps')
median(datf$totalsteps)
mean(datf$totalsteps)
```

#Are there differences in activity patterns between weekdays and weekends?

Revalue each date as "weekend" or "weekday"

```{r}
library(plyr)
datf <- dat
datf$steps[is.na(datf$steps)] <- mean(dat$steps,na.rm= TRUE)

datf$date <- weekdays(as.Date(datf$date))
datf$date <- revalue(datf$date, c("Saturday" = "weekend"))
datf$date <- revalue(datf$date, c("Sunday" = "weekend"))
datf$date <- revalue(datf$date, c("Monday" = "weekday"))
datf$date <- revalue(datf$date, c("Tuesday" = "weekday"))
datf$date <- revalue(datf$date, c("Wednesday" = "weekday"))
datf$date <- revalue(datf$date, c("Thursday" = "weekday"))
datf$date <- revalue(datf$date, c("Friday" = "weekday"))
head(datf)
```

Create 2 plots with average steps taken across weekday and weekend
```{r}
datwkday <- subset(datf, date == "weekday")
datwkday <- aggregate(datwkday$steps, by = list(datwkday$interval), FUN = mean, na.rm = TRUE)
names(datwkday) <- c("interval","mean_steps")

datwkend <- subset(datf, date == "weekend")
datwkend <- aggregate(datwkend$steps, by = list(datwkend$interval), FUN = mean, na.rm = TRUE)
names(datwkend) <- c("interval","mean_steps")

par(mfrow=c(2, 1))
plot(datwkday$interval, datwkday$mean_steps, type = "l", xlab = 'interval', ylab = "avg_Steps", main = 'Weekday')
plot(datwkend$interval, datwkend$mean_steps, type = "l", xlab = 'interval', ylab = "avg_Steps", main = 'Weekend')
```